---
title: "P8105_HW2_if2282"
author: Ifrah Fayyaz
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

## Reading Mr. Trash Wheel Dataset:

```{r}
trash_wheel = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N408"
    ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(homes_powered = round(homes_powered, digits = 0))
```

Description:
The dataset for Mr. Trashwheel contains `r ncol(trash_wheel)` variables and `r nrow(trash_wheel)` observations. The median number of sports balls in a dumpster in 2017 were `r filter(trash_wheel, year == 2017) %>% pull(sports_balls) %>% median()`. 

## Reading Precipiation Datasets for 2018 and 2019:

```{r}
precipitation_2018 = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14"
  ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(year = "2018")

precipitation_2019 = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14"
  ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>%
  mutate(month = month.name[month])
```

Description:
The dataset of Precipitation contains `r ncol(precipitation)` variables and `r nrow(precipitation)` observations. The total precipitation in the year 2018 was `r filter(precipitation, year == 2018) %>% pull(total) %>% sum()` and `r filter(precipitation, year == 2019) %>% pull(total) %>% sum()` in 2019. 


```{r}

```


```{r}

```


```{r}

```
