---
title: "P8105_HW2_if2282"
author: Ifrah Fayyaz
output: github_document
---

```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

## Reading Mr. Trash Wheel Dataset:

```{r, message = FALSE}
trash_wheel = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N408"
    ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(sports_balls = round(sports_balls, digits = 0)) %>%
  mutate(homes_powered = round(homes_powered, digits = 0))
```

Description:
The dataset for Mr. Trashwheel contains `r ncol(trash_wheel)` variables and `r nrow(trash_wheel)` observations. The median number of sports balls in a dumpster in 2017 were `r filter(trash_wheel, year == 2017) %>% pull(sports_balls) %>% median()`. The mean weight of trash accumulated by Mr. Trashweel in 2014 was `r filter(trash_wheel, year == 2014) %>% pull(weight_tons) %>% mean()` tons.

## Reading Precipiation Datasets for 2018 and 2019:

```{r, message = FALSE}
precipitation_2018 = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14"
  ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(year = "2018")

precipitation_2019 = 
  read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14"
  ) %>%
  janitor::clean_names() %>%
  na.omit() %>%
  mutate(year = "2019")

precipitation =
  bind_rows(precipitation_2018, precipitation_2019) %>%
  mutate(month = month.name[month])
```

Description:
The dataset of Precipitation contains `r ncol(precipitation)` variables and `r nrow(precipitation)` observations. The total precipitation in the year 2018 was `r filter(precipitation, year == 2018) %>% pull(total) %>% sum()` and `r filter(precipitation, year == 2019) %>% pull(total) %>% sum()` in 2019. 


# Problem 2

```{r, message = FALSE}
pol_months = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  separate(mon, sep = '-', into = c("year", "month", "day")) %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  mutate(president = recode(prez_dem, "1" = "Democrat", "0" = "Republican")) %>%
  select(-prez_dem, -prez_gop, -day)

snp = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  mutate(date = as.Date(date, "%m/%d/%y")) %>%
  separate(date, sep = '-', into = c("year", "month", "day")) %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  select(-day)

unemploy = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
    pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment"
  ) %>%
  janitor::clean_names() %>%
  mutate(year = as.character(year))

five38 =
  inner_join(pol_months, snp, by = c("year" = "year","month" = "month")) 

fivethree8 = 
  inner_join(five38, unemploy, by = c("year" = "year","month" = "month"))
```

The pol_months dataset is related to the number of national politicians who are democratic or republican at any given time. The dataset contained `r nrow(pol_months)` observations. The snp dataset is related to Standard & Poor’s stock market index (S&P), often used as a representative measure of stock market as a whole. The dataset contained `r nrow(snp)` observations. Lastly, the unemploy dataset contains the unemployment rates of the years and months. The dataset contained `r nrow(unemploy)` observations.

The fivethree8 dataset was a merge of all the three datasets listed above. It had a total of  `r nrow(fivethree8)` observations and `r ncol(fivethree8)` variables: `r names(fivethree8)`. The years included in this dataset ranged from `r range(pull(fivethree8, year))`.


# Problem 3

Some data tidying at first:

```{r, message = FALSE}
pop_baby =
  read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(childs_first_name = str_to_title(childs_first_name)) %>%
  mutate(ethnicity = recode(ethnicity, "BLACK NON HISP" = "BLACK NON HISPANIC")) %>%
  mutate(ethnicity = recode(ethnicity, "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER")) %>%
  distinct(year_of_birth, gender, ethnicity, childs_first_name, .keep_all = TRUE)
```

Table for the popularity of the name Olivia: 

```{r, message = FALSE}
olivia = 
  pop_baby %>%
  filter(childs_first_name == "Olivia") %>%
  select(year_of_birth, ethnicity, rank) %>%
  arrange(year_of_birth)
 
olivia %>%  
   pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>%
  knitr::kable()
```


Table for the most popular of the male name for different ethnicity over time: 

```{r, message = FALSE}
pop_male = 
  pop_baby %>%
  filter(gender == "MALE") %>%
  filter(rank == 1) %>%
  select(year_of_birth, childs_first_name, ethnicity) %>%
  arrange(year_of_birth)

pop_male %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>%
  knitr::kable()
```

Showing scatter plot for male, white non-Hispanic children born in 2016:

```{r, message = FALSE}
sc_plot = 
  pop_baby %>%
  filter(gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == 2016) %>%
  select(childs_first_name, count, rank) 

sc_plot %>%
  ggplot(aes(x = count, y = rank)) +
  geom_point(color = "purple") + 
  labs(
    title = "Count versus Rank plot for White non-Hispanic Males, 2016",
    x = "Count",
    y = "Rank"
  ) 
```


